Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2017-12-29 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Correction

 Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers~\citep{howison_2000}. During the displacement of a fluid by a less viscous one, the expected uniform front \citep{buckley_1942,sheldon_1959} is perturbed leading to an uneven front with elongations of the outside edge of the fluid interface (Fig.~\ref{fig:simple_case}), often referred as \textit{fingers}. In homogeneous domains, fingers start to develop when the surface tension acting on the interface between the fluids exert an opposite force towards the change of shape of the interface. The interface becomes unstable and collapses, taking a curved shape. \red{(KOSTAS: IS/ARE THERE ANY REFERENCE FOR THIS FINGERING FORMATION MECHANISMS?')}

\medskip 
In heterogeneous domains, such instability may be triggered by permeability differences across regions as demonstrated in Section~\ref{section:results_homo_hete}. The higher the velocity of the low viscosity fluid, the less wide (tip-splitting behaviour) the finger is. Pressure differences acting on the interface produces a net pressure force,      
\begin{equation} 
\Delta P= - \gamma \; \nabla \; \cdot \; \hat{n}. 
\label{eq:pressure_dif} 
\end{equation}

@@

\medskip
As demonstrated by \citet{habermann_1960}, \textit{mobility ratio} (MR) is a key-parameter to assess fluid displacement and is defined as the ratio of mobility of the displacing (fluid $i$) to that of the displaced fluid ($j$),
\begin{equation}
 \text{MR} \; = \frac{\mathcal{K}_{ri} \; \mu_{j}}{\mathcal{K}_{rj} \; \mu_{i}}. 
\label{eq:MR}
\end{equation}

\noindent MR is a function not only of fluids' viscosity and saturation, but also of the parameterised relative permeability, $\mathcal{K}_{r\alpha}$, which is often expressed as a function of the residual and maximum phase saturation prescribed in the pore rock matrix. In the simulations conducted for this work, the modified Brooks-Corey model was used \citep{alpak_1999,brooks_1964},

\begin{eqnarray}
  \mathcal{K}_{rw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rw}\left[\frc{S_{w}-S_{w,irr}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{w}}, \label{Eqn:CoreyBrooks1}\\
  \mathcal{K}_{rnw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rnw}\left[\frc{S_{nw}-S_{nw,r}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{nw}}, \label{Eqn:CoreyBrooks2}
\end{eqnarray}

>>>>

 Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers~\citep{howison_2000}. During the displacement of a fluid by a less viscous one, the expected uniform front \citep{buckley_1942,sheldon_1959} is perturbed leading to an uneven front with elongations of the outside edge of the fluid interface (Fig.~\ref{fig:simple_case}), often referred as \textit{fingers}. In homogeneous domains, fingers start to develop when the surface tension acting on the interface between the fluids exert an opposite force towards the change of shape of the interface. The interface becomes unstable and collapses, taking a curved shape. \red{(KOSTAS: IS/ARE THERE ANY REFERENCE FOR THIS FINGERING FORMATION MECHANISMS?)}

\medskip 
In heterogeneous domains, such instability may be triggered by permeability differences across regions as demonstrated in Section~\ref{section:results_homo_hete}. The higher the velocity of the low viscosity fluid, the less wide (tip-splitting behaviour) the finger is. Pressure differences acting on the interface produces a net pressure force,      
\begin{equation} 
\Delta P= - \gamma \nabla\cdot\hat{n}. 
\label{eq:pressure_dif} 
\end{equation}

@@

\medskip
As demonstrated by \citet{habermann_1960} \citep[see also][]{budek_2017}, mobility ratio (MR) is a key-parameter to assess fluid displacement and is defined as the ratio of mobility of the displacing (fluid $i$) to that of the displaced fluid ($j$),
\begin{equation}
 \text{MR} \; = \frac{\mathcal{K}_{ri} \mu_{j}}{\mathcal{K}_{rj} \mu_{i}}. 
\label{eq:MR}
\end{equation}
MR is a function not only of fluids' viscosity and saturation, but also of the parameterised relative permeability, $\mathcal{K}_{r\alpha}$, which is often expressed as a function of local, residual and maximum phase saturations prescribed in the pore rock matrix. In the simulations conducted for this work, the modified Brooks-Corey model was used \citep{alpak_1999,brooks_1964},
\begin{eqnarray}
  \mathcal{K}_{rw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rw}\left[\frc{S_{w}-S_{w,irr}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{w}}, \label{Eqn:CoreyBrooks1}\\
  \mathcal{K}_{rnw}\left(S_{w}\right) &=& \mathcal{K}^{\circ}_{rnw}\left[\frc{S_{nw}-S_{nw,r}}{1-S_{w,irr}-S_{nw,r}}\right]^{n_{nw}}, \label{Eqn:CoreyBrooks2}
\end{eqnarray}
